One page description

### 2016-10-03 publications/Journal/6-allocation.tex
Small changes to section 6

Although the synchronization protocol is one of the defining factors in simulation performance, model allocation has a significant impact on which protocol is ideal.
Depending on the model structure, and how it is mapped to the different kernels, it might not even make sense to parallelize at all.
Indeed, if the model is distributed such that frequent communication is necessary between kernels, parallelism is naturally reduced.
This brings us to the topic of model allocation, as also implemented in \textit{PythonPDEVS}~\cite{PythonPDEVS2}.

@@

The PHOLDTree model is similar in structure to models of gossiping in social networks~\cite{Gossip}.
The lookahead of an atomic node is the minimally allowed $\epsilon$, indicating uncertainty, as is often the case in realistic models.
We demonstrate the importance of allocation by comparing performance of a breadth-first versus a depth-first scheme.
Both options are automated ways of allocation that are independent of the model.

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtree.pdf}
    %TODO depth 1 but 3 levels deep?
    \caption{PHOLDTree model for depth 1 and width 2.}
    \label{fig:PHOLDTree_model}
\end{figure}

@@

\subsubsection{Sequential Simulation}
Since \textit{adevs} does not use direct connection, we expect a noticable performance difference between \textit{dxex} and \textit{adevs}.
This is shown in Figure~\ref{fig:PHOLDtree_seq_dn_benchmark}, where the fanout ($n$) determines the performance penalty \textit{adevs} suffers compared to \textit{dxex}.
Profiling indeed indicates that an increase in width per subtree ($n$) leads to higher overheads in \textit{adevs} due to the lack of direct connection.
\textit{Dxex} uses direct connection, making it independent on fanout.
Performance of \textit{dxex} is, in this model, only dependent on the number of models.
Slight deviations can still be seen, though, caused by the initialization overhead of direct connection.
Both \textit{adevs} and \textit{dxex} scale linearly in the number of atomic models.

@@

@@

@@

\begin{figure}
   \center
   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_BF.pdf}
   \caption{PHOLDTree model breadth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_bfs}
\end{figure}
\begin{figure}
   \center
   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_DF.pdf}
   \caption{PHOLDTree model depth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_dfs}
\end{figure}

@@

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4BFS.eps}
    \caption{Visualization of a PHOLDTree simulation with breadth first allocation and 4 kernels.}
    \label{fig:pholdtree_visualize_parBFS}
\end{figure}
\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4DFS.eps}
    \caption{Visualization of a PHOLDTree simulation with depth first allocation and 4 kernels.}
    \label{fig:pholdtree_visualize_parDFS}
\end{figure}

@@

%The effect is due to the topology of the kernels, in BFS the risk on reverts is higher. Optimistic is less influenced but will get a lower speedup in the good case.
Interestingly, we see that optimistic synchronization is less influenced by the allocation than conservative synchronization.
This is likely caused due to the lower number of connections to take into account in conservative synchronization.
Whereas conservative synchronization needs to take into account even scarcely used connections, optimistic synchronization does not.
The same is true in the opposite direction, though, where optimistic synchronization is slower when a good allocation is chosen.
Conservative synchronization will then be able to make better estimates, whereas optimistic synchronization does not make estimations.

>>>>

Although the synchronization protocol is one of the defining factors in simulation performance, model allocation has a significant impact on which protocol is ideal.
Depending on the model structure, and how it is mapped to the different kernels, it might not even make sense to parallelize at all.
Indeed, if the model is allocated such that frequent communication is necessary between kernels, parallelism is naturally reduced.
This brings us to the topic of model allocation, as also implemented in \textit{PythonPDEVS}~\cite{PythonPDEVS2}.

@@

The PHOLDTree model is similar in structure to models of gossiping in social networks~\cite{Gossip}.
The lookahead of an atomic node is the minimally allowed $\epsilon$, as is often the case in realistic models (\textit{e.g.}, if it is unknown what the lookahead might be).
We demonstrate the importance of allocation by comparing performance of a breadth-first versus a depth-first scheme.
Both options are automated ways of allocation that are independent of the model.

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtree.pdf}
    \caption{PHOLDTree model for depth 3 and width 2.}
    \label{fig:PHOLDTree_model}
\end{figure}

@@

\subsubsection{Sequential Simulation}
Since \textit{adevs} does not use direct connection, we expect a noticeable performance difference between \textit{dxex} and \textit{adevs}.
This is shown in Figure~\ref{fig:PHOLDtree_seq_dn_benchmark}, where the fanout ($n$) determines the performance penalty \textit{adevs} suffers compared to \textit{dxex}.
Profiling indeed indicates that an increase in width per subtree ($n$) leads to higher overheads in \textit{adevs} due to the lack of direct connection.
\textit{Dxex} uses direct connection, making it independent of fanout.
Performance of \textit{dxex} is, in this model, only dependent on the number of models.
Slight deviations can still be seen, though, caused by the initialization overhead of direct connection.
Both \textit{adevs} and \textit{dxex} scale linearly in the number of atomic models.

@@

\begin{figure}
   \center
   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_BF.pdf}
   \caption{PHOLDTree model breadth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_bfs}
\end{figure}

@@

\begin{figure}
   \center
   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_DF.pdf}
   \caption{PHOLDTree model depth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_dfs}
\end{figure}

@@

@@

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4BFS.eps}
    \caption{Visualization of a simulation of the model in Figure~\ref{fig:PholdTree_model_bfs}.}
    \label{fig:pholdtree_visualize_parBFS}
\end{figure}

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4DFS.eps}
    \caption{Visualization of a simulation of the model in Figure~\ref{fig:PholdTree_model_dfs}.}
    \label{fig:pholdtree_visualize_parDFS}
\end{figure}

@@

%The effect is due to the topology of the kernels, in BFS the risk on reverts is higher. Optimistic is less influenced but will get a lower speedup in the good case.
Interestingly, we see that optimistic synchronization is less influenced by the allocation than conservative synchronization.
This is likely caused by the lower number of connections to take into account in conservative synchronization.
Whereas conservative synchronization needs to take into account even scarcely used connections, optimistic synchronization does not.
The same is true in the opposite direction, though, where optimistic synchronization is slower when a good allocation is chosen.
Conservative synchronization will then be able to make better estimates, whereas optimistic synchronization does not make estimations.
